---
title: "Installing RHugin on Microsoft Windows"
author: "Kjell Konis"
date: "January 26, 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


***

Read the *Preliminaries* section even if you already have R, Rtools, and the Hugin Decision Engine installed. You probably will need to install the *Microsoft Visual C++ Redistributable Package (x64)*.  These instructions have only been tested with 64 bit versions of R and the HDE.


***

## Preliminaries

The RHugin package provides an R API for the Hugin Decision Engine (HDE): both the HDE and R must be installed on the computer in order for the RHugin package to function.


### Install HUGIN and the HUGIN Decision Engine (HDE)

The Hugin Decision Engine is commercial software available from HUGIN EXPERT.

> [http://www.hugin.com](http://www.hugin.com)

The products *HUGIN Developer* (commercial use) and *HUGIN Researcher* (academic use) include the Hugin Decision Engine. The HDE is also included with the free demo *HUGIN Lite* (the only permitted use of the free demo is to evaluate the HUGIN software).


### Install the Microsoft Visual C++ Redistributable Package (x64)



### Install R

R is available from the CRAN repository.

> [https://cloud.r-project.org/](https://cloud.r-project.org)

The installer for the most recent version of R can be found by following *Download R for Windows > base*.


### Install Rtools

Since RHugin is distributed as a source package you will also need to install Rtools. The Rtools installer can be found by following *Download R for Windows > Rtools*.  The most recent version is at the top of the list.

Once these three pieces of software are installed, start R and proceed to the next section.


***

## RHugin Package Depedencies

```{r uninstallBiocPackages, echo = FALSE, message = FALSE, results = "hide"}
p <- intersect(pkgs <- installed.packages()[, 1],
               c("Rgraphviz", "graph", "BiocGenerics", "BiocInstaller"))
remove.packages(p)
```



```{r installRgraphviz, message = FALSE, results = "hide"}
source("https://bioconductor.org/biocLite.R")
biocLite(c("graph", "Rgraphviz"))
```


## Install the RHugin Package

```{r unInstallRHugin}
if("RHugin" %in% pkgs) remove.packages("RHugin")
```


```{r installRHugin}
install.packages("RHugin", repos = "http://rhugin.r-forge.r-project.org",
                 type = "source", INSTALL_opts="--no-multiarch")
```



