## `RHugin` Install Instructions for Mac OS X
Last updated by Kjell Konis on 2017-02-02.


***

### Preliminaries

The `RHugin` package provides an R API for the HUGIN Decision Engine (HDE): both the HDE and R must be installed on the computer in order for the `RHugin` package to function.  Since `RHugin` is distributed as a source package, the *command line developer tools* are also required.


#### Install HUGIN and the HUGIN Decision Engine (HDE)

The HUGIN Decision Engine is commercial software available from HUGIN EXPERT.

> [http://www.hugin.com](http://www.hugin.com)

The products *HUGIN Developer* (commercial use) and *HUGIN Researcher* (academic use) include the HUGIN Decision Engine. The HDE is also included with the free demo *HUGIN Lite* (the only permitted use of the free demo is to evaluate HUGIN). HUGIN is distributed on a disk image (i.e., a `.dmg` file). Mount the disk image by double clicking on it, navigate to the mounted disk, and install the HDE by dragging the HDE folder and dropping it onto the *Applications* alias. You may also want to install the other components (HUGIN GUI, documentation, etc.) but these are not required for `RHugin`.


#### Install R

Just in case you don't already have R installed... R is available from the CRAN repository.

> [https://cloud.r-project.org/](https://cloud.r-project.org)

The installer for the most recent version of R can be found by following the *Download R for (Mac) OS X* link.


#### Install *Command Line Developer Tools*

Start R and run the following command to install the *command line developer tools*.

```{r installCLDT, eval = FALSE}
system2("xcode-select", "--install")
```

A dialog box will pop up and prompt you to install the command line developer tools. Click *Install* (note: this command must be run by a user with administrative privileges).

If you get the following error

```{r CLDTAlreadyInstalled, echo = FALSE, comment = NA}
cat("xcode-select: error: command line tools are already installed, use \"Software Update\" to install updates")
```

the command line developer tools are already installed and you could have skipped this section.


***

### RHugin Package Depedencies

```{r uninstallBiocPackages, echo = FALSE, message = FALSE, results = "hide"}
p <- intersect(pkgs <- installed.packages()[, 1],
               c("Rgraphviz", "graph", "BiocGenerics", "BiocInstaller"))
remove.packages(p)
```

The RHugin package has dependencies on the `graph` and `Rgraphviz` packages.  These packages are part of Bioconductor and can be installed by running the following two R commands.

```{r installRgraphviz, message = FALSE, results = "hide"}
source("https://bioconductor.org/biocLite.R")
biocLite(c("graph", "Rgraphviz"))
```


***

### Install the RHugin Package

```{r unInstallRHugin, echo = FALSE, result = "hide", message = FALSE}
if("RHugin" %in% pkgs) remove.packages("RHugin")
```

First, set the `HUGINHOME` environment variable so that `RHugin` can find the HUGIN Decision Engine (HDE).  The value (i.e., the `"/Applications/HDE8.4-x64"` in the code below) should be the full path to the HDE folder that you installed in the *Preliminaries* section.

```{r setHUGINHOME}
Sys.setenv(HUGINHOME = "/Applications/HDE8.4-x64")
```

Then, install the `RHugin` package from `RHugin`'s R-Forge repository with the following R command.

```{r installRHugin, result = "hide", message = FALSE}
install.packages("RHugin", repos = "http://rhugin.r-forge.r-project.org",
                 type = "source", INSTALL_opts="--no-multiarch")
```


***

### Quick Test

To check that everything went as planned run the following code to plot one of the example networks provided by HUGIN.

```{r testRHuginInstallation, fig.align = "center"}
library(RHugin)
demo("asia", echo = FALSE)
plot(asia)
```

***

Happy HUGINing with `RHugin`!


***
***
***

Copyright 2017 Kjell Konis








